DOpElib: Differential Equations and Optimization Environment; A Goal Oriented Software Library for Solving PDEs and Optimization Problems

### 2012-11-28 Paper_May_2012/main.tex
Antworten auf CGs Vorschlaege und am Anfang zwei neue Todos

\institute{Christian Goll \at
              Institute of Applied Mathematics, University of Heidelberg
              \\Im Neuenheimer Feld 293/294, 69120 Heidelberg, Germany\\
              Tel.: +49 622-154-6154\\
              \email{cgoll@iwr.uni-heidelberg.de}           %  \\
           \and
Thomas Wick \at
              The Institute for Computational Engineering and Sciences,\\
The University of Texas at Austin,\\
Austin, Texas 78712, USA \\
              Tel.: +1 512-672-7763\\
              \email{twick@ices.utexas.edu}   
 \and
W. Wollner \at
              Universit\"at Hamburg\\
              Department of Mathematics\\
              \email{winnifried.wollner@math.uni-hamburg.de}   
}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
\todo{Wird die Webseite genannt? ist das sinnvoll dies zu tun?}\todocg{Seite 2, vorletzter Absatz. Da wird DOpElib referenziert, und da steht dann auch die Webseite.}
In this article, we describe the software library 
\textit{Differential Equations and Optimization Environment} (\dope{}).
The main feature of \dope{} is that it provides a unified interface to high level algorithms 
such as time-stepping methods, nonlinear solvers and optimization routines. This structure ensures 
that first of all the user is required to write only those sections of code that are specific to 
the considered problem. Second, the exchange of parts of the used routines is possible 
with only a few lines of code to change instead of large reimplementations.
The article illustrates the design principles and various features
of \dope{} and provides some 
numerical results as demonstration for the versatility of the software.
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{introduction}
Today, there exists a broad variety of software libraries
that provide numerical tools for solving partial differential
equations (PDEs). Prominent examples are 
\deal{} \citep{dealnew}, dune \citep{dune}, 
FEniCS \citep{fenics}
or commercial solvers as for instance ANSYS (with its Fluent package),
COMSOL, or simufact forming.
Finally, a lot of work is (still) done in Matlab.
In fact, these libraries provide very useful tools 
for a huge variety of applications.

@@

In the next section we will now sketch the way a user needs to implement the
user-dependent data as the local cell integrals and the initialization of the
objects described above.
\subsubsection{Problem dependent implementation of stationary PDE problems}
To solve a stationary PDE problem, the user has to provide some code which can roughly be divided into the following main sections:
\begin{description}
\item[\textit{Discretization}]\todocg{Hier scheint etwas verlorengegangen zu
    sein, der Satz macht so keinen Sinn:} 
\todo{Vorschlag: Specification of triangulation issues such as finite elements
  and quadrature rules}  
  \todocg{Gegenvorschlag: Specification of discretization issues...}
Specify what kind of triangulation:
in most cases this is the
finite element and quadrature rules. Furthermore, other
things might be adjusted as well.
\item[\textit{Problem-data}] \todo{Definition of } Define the weak form, the corresponding matrices, the
  functionals $J$ as well as the boundary conditions. With exception of the boundary conditions, we have to specify all this only cell-wise, \dope{} handles the rest for us.
\item[\textit{Solvers}] Choice of a (non-)linear solver. 
\item[\textit{Additional features}] ParameterReader to read in the run-time parameters
  such as material coefficients or adjustments for the solvers like tolerance
  for Newton's method. 
\end{description}

@@

\subsubsection{\texttt{localfunctional.h}}
The implementation of the boundary-functional $J$ follows the outline of the
computation of the cell-residual in \texttt{localpde.h}. 
%The main difference, and common to all functionals is that we only want to evaluate a functional 
\todocg{Ist hier was verrutscht? der Satz passt im
  Moment nicht so wirklich, der Teil mit dem Vektor ist irgendwie
  verlorengegangen?}. 
\todo{Ich hatte Deine Anmerkung so verstanden, als ob Du alles nach dem i.e.,
  geloescht 
haben wolltest. Im main TW steht noch das Original}
\todocg{Ich habe den Rest auskommentiert. Mein Vorschlag waere: The main difference is that as we only want to evaluate a functional, i.e. a number, and not a residual-vector, we just loop over the quadrature points and not over the degrees of freedom on the face. }
%Thus, we just loop over the quadrature points and not over the degrees of freedom on the face.
In the present situation, we want to evaluate a functional 
that consists of a boundary integral. Thus we need to evaluate 
integrals on faces, those who lie on $\Gamma_{circ}$, instead of integrals over
elements. 
This face contributions are calculated in \texttt{BoundaryValue} and added to 
sum up to the value of $J$. 
\begin{lstlisting}
 double
 BoundaryValue(const FaceDataContainer& fdc)
  {
    //same as before, get n_q_points etc.
    const auto& face_values = cdc.GetFEFaceValuesState();
    fdc.GetFaceValuesState("state", ufacevalues);
    fdc.GetFaceGradsState("state", ufacegrads);
\end{lstlisting}
Notice that we get here the values of the computed solution by giving the string "state".
\begin{lstlisting}
    for(unsigned int q_point = 0; 
        q_point < n_q_points; 
        q_point++)
      erg -= ( _nu * ufacegrads[q_point][0]
               * face_values.normal_vector(q_point)
               + //other terms        
             ) * face_values.JxW(q_point);
    return erg;
  }
\end{lstlisting}

>>>>

\institute{Christian Goll \at
              Institute of Applied Mathematics, University of Heidelberg
              \\Im Neuenheimer Feld 293/294, 69120 Heidelberg, Germany\\
              Tel.: +49 622-154-6154\\
              \email{cgoll@iwr.uni-heidelberg.de}           %  \\
           \and
Thomas Wick \at
              The Institute for Computational Engineering and Sciences,\\
The University of Texas at Austin,\\
Austin, Texas 78712, USA \\
              %Tel.: +1 512-672-7763\\
              \email{twick@ices.utexas.edu}   
 \and
W. Wollner \at
              Universit\"at Hamburg\\
              Department of Mathematics\\
              \email{winnifried.wollner@math.uni-hamburg.de}   
}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
\todo{Ich finde die Adressen uneinheitlich. CG und ich koennten die
  Telefonnummern wegstreichen, um uns an Winni anzupassen. Und Winni
keonnte Hausnummer oder Postleitzahl hinzufuegen. Immerhin ist 
das Standart in (fast) allen Journals}
In this article, we describe the software library 
\textit{Differential Equations and Optimization Environment} (\dope{}).
The main feature of \dope{} is that it provides a unified interface to high level algorithms 
such as time-stepping methods, nonlinear solvers and optimization routines. This structure ensures 
that first of all the user is required to write only those sections of code that are specific to 
the considered problem. Second, the exchange of parts of the used routines is possible 
with only a few lines of code to change instead of large reimplementations.
The article illustrates the design principles and various features
of \dope{} and provides some 
numerical results as demonstration for the versatility of the software.
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{introduction}
\todo{Reihenfolge bei der Erwaehnung von Dopelib mit vollem Namen und 
des Anfangs in 2010 passt nicht ueberein. 2010 wird am Ende erwahent und der
volle Name von Dope wird erst genannt nachdem wir bereits die Abkuerzung
benutzt haben. Das geht so nicht!}
Today, there exists a broad variety of software libraries
that provide numerical tools for solving partial differential
equations (PDEs). Prominent examples are 
\deal{} \citep{dealnew}, dune \citep{dune}, 
FEniCS \citep{fenics}
or commercial solvers as for instance ANSYS (with its Fluent package),
COMSOL, or simufact forming.
Finally, a lot of work is (still) done in Matlab.
In fact, these libraries provide very useful tools 
for a huge variety of applications.

@@

In the next section we will now sketch the way a user needs to implement the
user-dependent data as the local cell integrals and the initialization of the
objects described above.
\subsubsection{Problem dependent implementation of stationary PDE problems}
To solve a stationary PDE problem, the user has to provide some code which can roughly be divided into the following main sections:
\begin{description}
\item[\textit{Discretization}]\todocg{Hier scheint etwas verlorengegangen zu
    sein, der Satz macht so keinen Sinn:} 
\todo{Vorschlag: Specification of triangulation issues such as finite elements
  and quadrature rules}  
  \todocg{Gegenvorschlag: Specification of discretization issues...}
\todo{Finde ich gut: Koennen wir machen}
Specify what kind of triangulation:
in most cases this is the
finite element and quadrature rules. Furthermore, other
things might be adjusted as well.
\item[\textit{Problem-data}] \todo{Definition of } Define the weak form, the corresponding matrices, the
  functionals $J$ as well as the boundary conditions. With exception of the boundary conditions, we have to specify all this only cell-wise, \dope{} handles the rest for us.
\item[\textit{Solvers}] Choice of a (non-)linear solver. 
\item[\textit{Additional features}] ParameterReader to read in the run-time parameters
  such as material coefficients or adjustments for the solvers like tolerance
  for Newton's method. 
\end{description}

@@

\subsubsection{\texttt{localfunctional.h}}
The implementation of the boundary-functional $J$ follows the outline of the
computation of the cell-residual in \texttt{localpde.h}. 
%The main difference, and common to all functionals is that we only want to evaluate a functional 
\todocg{Ist hier was verrutscht? der Satz passt im
  Moment nicht so wirklich, der Teil mit dem Vektor ist irgendwie
  verlorengegangen?}. 
\todo{Ich hatte Deine Anmerkung so verstanden, als ob Du alles nach dem i.e.,
  geloescht 
haben wolltest. Im main TW steht noch das Original}
\todocg{Ich habe den Rest auskommentiert. Mein Vorschlag waere: The main
  difference is that as we only want to evaluate a functional, i.e. a number,
  and not a residual-vector, we just loop over the quadrature points and not
  over the degrees of freedom on the face. }
\todo{Finde ich gut. Kannst Du so machen!}
%Thus, we just loop over the quadrature points and not over the degrees of freedom on the face.
In the present situation, we want to evaluate a functional 
that consists of a boundary integral. Thus we need to evaluate 
integrals on faces, those who lie on $\Gamma_{circ}$, instead of integrals over
elements. 
This face contributions are calculated in \texttt{BoundaryValue} and added to 
sum up to the value of $J$. 
\begin{lstlisting}
 double
 BoundaryValue(const FaceDataContainer& fdc)
  {
    //same as before, get n_q_points etc.
    const auto& face_values = cdc.GetFEFaceValuesState();
    fdc.GetFaceValuesState("state", ufacevalues);
    fdc.GetFaceGradsState("state", ufacegrads);
\end{lstlisting}
Notice that we get here the values of the computed solution by giving the string "state".
\begin{lstlisting}
    for(unsigned int q_point = 0; 
        q_point < n_q_points; 
        q_point++)
      erg -= ( _nu * ufacegrads[q_point][0]
               * face_values.normal_vector(q_point)
               + //other terms        
             ) * face_values.JxW(q_point);
    return erg;
  }
\end{lstlisting}
